### 2016-09-20 writing/tables/old/1_robust.tex

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular*}{\hsize}{@{\hskip\tabcolsep\extracolsep\fill}l*{4}{c}}
\midrule \vspace{5mm} \underline{\textbf{Panel C : Text}}\hphantom{ations}\vspace{5mm} ($\mu$=5675)\\ label
\emph{treat X post}&     -0.0424         &      2284.2         &      5337.6         &      0.0572         \\
            &    (0.0934)         &    (2256.4)         &    (1469.0)\sym{***}&    (0.0950)         \\
\midrule
Player FE   &         Yes         &         Yes         &         Yes         &         Yes         \\
Year FE     &         Yes         &         Yes         &         Yes         &         Yes         \\
N           &        7033         &        4966         &        7033         &        5291         \\
Clusters    &         541         &         382         &         541         &         407         \\
\bottomrule
\end{tabular*}
}