### 2016-09-20 writing/tables/old/appendix_alternate.tex

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular*}{\hsize}{@{\hskip\tabcolsep\extracolsep\fill}l*{6}{c}}
\toprule
            &\multicolumn{3}{c}{Sample A}                                     &\multicolumn{3}{c}{Sample B}                                     \\\cmidrule(lr){2-4}\cmidrule(lr){5-7}
            &\multicolumn{1}{c}{Citations}&\multicolumn{1}{c}{Images}&\multicolumn{1}{c}{Text}&\multicolumn{1}{c}{Citations}&\multicolumn{1}{c}{Images}&\multicolumn{1}{c}{Text}\\
\midrule
\emph{out-of-copy X post}&    -0.00835         &      0.0911         &     -0.0994         &      0.0631         &      0.0631         &      0.0631         \\
            &     (0.266)         &     (0.108)         &     (0.195)         &    (0.0398)         &    (0.0398)         &    (0.0398)         \\
\midrule
Player FE   &         Yes         &         Yes         &         Yes         &         Yes         &         Yes         &         Yes         \\
Time FE     &        Year         &        Year         &        Year         &        Year         &        Year         &        Year         \\
adj. $R^2$  &       0.320         &      0.0314         &       0.385         &      0.0350         &      0.0350         &      0.0350         \\
N           &         240         &         240         &         240         &        3246         &        3246         &        3246         \\
\bottomrule
\end{tabular*}
}