### 2016-09-20 writing/tables/old/appendix_alternate_2012_2004_2007.tex

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular*}{\hsize}{@{\hskip\tabcolsep\extracolsep\fill}l*{6}{c}}
\toprule
            &\multicolumn{3}{c}{Sample A}                                     &\multicolumn{3}{c}{Sample B}                                     \\\cmidrule(lr){2-4}\cmidrule(lr){5-7}
            &\multicolumn{1}{c}{Citations}&\multicolumn{1}{c}{Images}&\multicolumn{1}{c}{Text}&\multicolumn{1}{c}{Citations}&\multicolumn{1}{c}{Images}&\multicolumn{1}{c}{Text}\\
\midrule
\emph{out-of-copy X post}&       3.924         &       3.647         &       0.277         &       0.150         &       0.150         &       0.150         \\
            &     (1.207)\sym{***}&     (0.729)\sym{***}&     (0.681)         &    (0.0782)\sym{*}  &    (0.0782)\sym{*}  &    (0.0782)\sym{*}  \\
\midrule
Player FE   &         Yes         &         Yes         &         Yes         &         Yes         &         Yes         &         Yes         \\
Time FE     &        Year         &        Year         &        Year         &        Year         &        Year         &        Year         \\
adj. $R^2$  &       0.686         &       0.328         &       0.812         &      0.0731         &      0.0731         &      0.0731         \\
N           &         360         &         360         &         360         &        4869         &        4869         &        4869         \\
\bottomrule
\end{tabular*}
}