### 2016-09-20 writing/tables/old/reg_dd_combined.tex

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular*}{\hsize}{@{\hskip\tabcolsep\extracolsep\fill}l*{6}{c}}
\toprule
            &\multicolumn{3}{c}{Sample A}                                     &\multicolumn{3}{c}{Sample B}                                     \\\cmidrule(lr){2-4}\cmidrule(lr){5-7}
            &\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Log-Cites}&\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Log-Cites}\\
\midrule
\emph{out-of-copy X post}&       5.825         &       5.798         &       0.347         &       0.216         &       0.206         &      0.0774         \\
            &     (1.820)\sym{***}&     (1.797)\sym{***}&     (0.153)\sym{**} &    (0.0592)\sym{***}&     (0.109)\sym{*}  &    (0.0370)\sym{**} \\
\midrule
Unit of Obs. FE&          No         &         Yes         &         Yes         &          No         &         Yes         &         Yes         \\
Year FE     &        Year         &        Year         &        Year         &         Yes         &         Yes         &         Yes         \\
adj. $R^2$  &       0.645         &       0.710         &       0.912         &      0.0565         &      0.0779         &       0.102         \\
N           &         360         &         360         &         360         &        6492         &        6492         &        6492         \\
\bottomrule
\end{tabular*}
}