### 2016-09-20 writing/tables/old/reg_dd_combined_2006_2010_2009.tex

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular*}{\hsize}{@{\hskip\tabcolsep\extracolsep\fill}l*{6}{c}}
\toprule
            &\multicolumn{3}{c}{Sample A}                                     &\multicolumn{3}{c}{Sample B}                                     \\\cmidrule(lr){2-4}\cmidrule(lr){5-7}
            &\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Log-Cites}&\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Log-Cites}\\
\midrule
\emph{out-of-copy X post}&       0.975         &       0.917         &      0.0117         &       0.177         &       0.152         &      0.0540         \\
            &     (0.972)         &     (0.945)         &     (0.157)         &    (0.0745)\sym{**} &    (0.0914)\sym{*}  &    (0.0316)\sym{*}  \\
\midrule
Unit of Obs. FE&          No         &         Yes         &         Yes         &          No         &         Yes         &         Yes         \\
Year FE     &         Yes         &         Yes         &         Yes         &         Yes         &         Yes         &         Yes         \\
adj. $R^2$  &       0.489         &       0.580         &       0.829         &      0.0381         &      0.0595         &      0.0757         \\
N           &         200         &         200         &         200         &        2705         &        2705         &        2705         \\
\bottomrule
\end{tabular*}
}