### 2016-09-20 writing/tables/old/reg_dd_compare.tex

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular*}{\hsize}{@{\hskip\tabcolsep\extracolsep\fill}l*{6}{c}}
\toprule
            &\multicolumn{3}{c}{Images}                                       &\multicolumn{3}{c}{Text}                                         \\\cmidrule(lr){2-4}\cmidrule(lr){5-7}
            &\multicolumn{1}{c}{OLS}&\multicolumn{1}{c}{OLS}&\multicolumn{1}{c}{Log-OLS}&\multicolumn{1}{c}{OLS}&\multicolumn{1}{c}{OLS}&\multicolumn{1}{c}{Log-OLS}\\
\midrule
\emph{out-of-copy X post}&       5.471         &       5.471         &       1.178         &       0.354         &       0.328         &     0.00783         \\
            &     (1.093)\sym{***}&     (1.093)\sym{***}&     (0.152)\sym{***}&     (1.023)         &     (1.009)         &     (0.141)         \\
\midrule
Publication-Year FE&          No         &         Yes         &         Yes         &          No         &         Yes         &         Yes         \\
Year FE     &         Yes         &         Yes         &         Yes         &         Yes         &         Yes         &         Yes         \\
adj. $R^2$  &       0.415         &       0.417         &       0.574         &       0.737         &       0.812         &       0.918         \\
N           &         360         &         360         &         360         &         360         &         360         &         360         \\
\bottomrule
\end{tabular*}
}