### 2016-09-20 writing/tables/old/ttest_0.tex

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular}{l*{1}{cccc}}
\toprule
                    &\multicolumn{4}{c}{}                               \\
                    &\textbf{(1)out-of-copy $\bar{y}$}&\textbf{(2)in-copy $\bar{y}$}&\textbf{(3)diff}&\textbf{(4)p-val}\\
\midrule
\emph{Number of Citations to Baseball Digest}&           0&           0&           0&           .\\
\emph{Number of Images}&       0.568&       0.411&       0.157&        0.17\\
\emph{Number of Words of Text}&      1340.4&      1189.4&       151.0&        0.38\\
\emph{Average Monthly Traffic}&       134.4&       175.8&      -41.40&        0.66\\
\bottomrule
\end{tabular}
}