### 2016-09-20 writing/tables/old/ttest_1.tex

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular}{l*{1}{cccc}}
\toprule
                    &\multicolumn{4}{c}{}                               \\
                    &\textbf{(1)out-of-copy $\bar{y}$}&\textbf{(2)in-copy $\bar{y}$}&\textbf{(3)diff}&\textbf{(4)p-val}\\
\midrule
\emph{Number of Citations to Baseball Digest}&       0.597&       0.334&       0.263&        0.03\\
\emph{Number of Images}&       1.888&       1.036&       0.853&        0.00\\
\emph{Number of Words of Text}&      2199.9&      1722.8&       477.1&        0.00\\
\emph{Average Monthly Traffic}&       190.3&       93.35&       96.94&        0.02\\
\bottomrule
\end{tabular}
}