### 2016-09-20 writing/tables/reg_dd_combined_2004_2009_2007.tex

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular*}{\hsize}{@{\hskip\tabcolsep\extracolsep\fill}l*{6}{c}}
\toprule
            &\multicolumn{3}{c}{Sample A}                                     &\multicolumn{3}{c}{Sample B}                                     \\\cmidrule(lr){2-4}\cmidrule(lr){5-7}
            &\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Log-Cites}&\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Log-Cites}\\
\midrule
\emph{out-of-copy X post}&      0.0493         &      0.0618         &      0.0311         &      0.0685         &      0.0631         &      0.0207         \\
            &     (0.300)         &     (0.271)         &    (0.0855)         &    (0.0323)\sym{**} &    (0.0398)         &    (0.0145)         \\
\midrule
Unit of Obs. FE&          No         &         Yes         &         Yes         &          No         &         Yes         &         Yes         \\
Year FE     &         Yes         &         Yes         &         Yes         &         Yes         &         Yes         &         Yes         \\
adj. $R^2$  &       0.256         &       0.319         &       0.525         &      0.0265         &      0.0350         &      0.0441         \\
N           &         240         &         240         &         240         &        3246         &        3246         &        3246         \\
\bottomrule
\end{tabular*}
}