### 2016-09-20 writing/tables/reg_dd_combined_2004_2012_2009.tex

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular*}{\hsize}{@{\hskip\tabcolsep\extracolsep\fill}l*{6}{c}}
\toprule
            &\multicolumn{3}{c}{Sample A}                                     &\multicolumn{3}{c}{Sample B}                                     \\\cmidrule(lr){2-4}\cmidrule(lr){5-7}
            &\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Log-Cites}&\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Log-Cites}\\
\midrule
\emph{out-of-copy X post}&       5.667         &       5.609         &       0.316         &       0.202         &       0.188         &      0.0690         \\
            &     (1.830)\sym{***}&     (1.809)\sym{***}&     (0.153)\sym{**} &    (0.0588)\sym{***}&     (0.107)\sym{*}  &    (0.0360)\sym{*}  \\
\midrule
Unit of Obs. FE&          No         &         Yes         &         Yes         &          No         &         Yes         &         Yes         \\
Year FE     &         Yes         &         Yes         &         Yes         &         Yes         &         Yes         &         Yes         \\
Page-Age FE &         ---         &         ---         &         ---         &         Yes         &         Yes         &         Yes         \\
adj. $R^2$  &       0.637         &       0.703         &       0.908         &      0.0592         &      0.0851         &       0.116         \\
N           &         360         &         360         &         360         &        4869         &        4869         &        4869         \\
\bottomrule
\end{tabular*}
}