### 2016-09-20 writing/tables/reg_dd_combined_2005_2011_2009.tex

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular*}{\hsize}{@{\hskip\tabcolsep\extracolsep\fill}l*{6}{c}}
\toprule
            &\multicolumn{3}{c}{Sample A}                                     &\multicolumn{3}{c}{Sample B}                                     \\\cmidrule(lr){2-4}\cmidrule(lr){5-7}
            &\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Log-Cites}&\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Cites}&\multicolumn{1}{c}{Log-Cites}\\
\midrule
\emph{out-of-copy X post}&       4.035         &       3.951         &       0.198         &       0.199         &       0.180         &      0.0655         \\
            &     (1.487)\sym{***}&     (1.459)\sym{***}&     (0.150)         &    (0.0657)\sym{***}&     (0.102)\sym{*}  &    (0.0345)\sym{*}  \\
\midrule
Unit of Obs. FE&          No         &         Yes         &         Yes         &          No         &         Yes         &         Yes         \\
Year FE     &         Yes         &         Yes         &         Yes         &         Yes         &         Yes         &         Yes         \\
adj. $R^2$  &       0.619         &       0.678         &       0.889         &      0.0444         &      0.0695         &      0.0898         \\
N           &         280         &         280         &         280         &        3787         &        3787         &        3787         \\
\bottomrule
\end{tabular*}
}