### 2016-09-20 writing/tables/reg_dd_compare_sampleb.tex

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular*}{\hsize}{@{\hskip\tabcolsep\extracolsep\fill}l*{6}{c}}
\toprule
            &\multicolumn{3}{c}{Images}                                       &\multicolumn{3}{c}{Text}                                         \\\cmidrule(lr){2-4}\cmidrule(lr){5-7}
            &\multicolumn{1}{c}{OLS}&\multicolumn{1}{c}{OLS}&\multicolumn{1}{c}{Log-OLS}&\multicolumn{1}{c}{OLS}&\multicolumn{1}{c}{OLS}&\multicolumn{1}{c}{Log-OLS}\\
\midrule
\emph{out-of-copy X post}&       0.477         &       0.337         &       0.109         &       0.466         &       0.314         &      -0.113         \\
            &     (0.102)\sym{***}&    (0.0819)\sym{***}&    (0.0272)\sym{***}&     (0.120)\sym{***}&    (0.0877)\sym{***}&    (0.0490)\sym{**} \\
\midrule
Player-Page FE&          No         &         Yes         &         Yes         &          No         &         Yes         &         Yes         \\
Year FE     &         Yes         &         Yes         &         Yes         &         Yes         &         Yes         &         Yes         \\
adj. $R^2$  &      0.0992         &       0.159         &       0.205         &       0.218         &       0.394         &       0.844         \\
N           &       13260         &       13260         &       13260         &       13260         &       13260         &       13260         \\
\bottomrule
\end{tabular*}
}